Repeatability of Pre-Flashover Fire Patterns on Gypsum Wall Board

### 2015-03-07 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Add .bib file for Madrzykowski thesis.

@@

Uncontrolled fires result in loss of life, property and can create an adverse economic impact on a community.  Findings from an archaeological site in South Africa, published in 2012, indicate that mankind has productively used fire for approximately 1 million years.[1]  This is 300 thousand years earlier than previously thought.  Despite mankind’s extensive history with fire, according to the World Health Organization, an estimated 265,000 people die each year as a result of burn injuries.[2]  Based on an estimated global population of 7.3 billion people, the death rate due to burns from all sources would be 3.6 per 100,000 populations.  The World Fire Statistics Center collects data from European, North American and a few Asia-Oceania countries.  Within this group of countries the fire deaths per 100,000 population range from 0.02 (Singapore) to 2.03 (Finland). [3] There are a significant number of factors that determine the variability of fire death rates between nations, socio- economic conditions being a leading factor.[2]

In the United States, the loss of life due to unwanted fire has been decreasing steadily since the 1970’s.  This is neglecting the increased loss of life as a result of the terrorist attack on September 11, 2001.  The estimated number of fire deaths in the U.S. in 1975 was 8,100 or 3.75 deaths per 100,000 population.[4]  In 2013, there were 3,240 civilian fire fatalities or 1.02 deaths per 100,000 population.[5]  During this same period, the number of fires in the United States has decreased from 2.5 million fires to 1.24 million fires.  There are many reasons that led to this decrease in loss of life due to fire in the United States.  An example would be research and data that enabled the development of product safety standards, improved building codes, improved fire safety systems, and fire prevention programs to name just a few.  None of this happened without the constant effort of agencies at the Federal, state, and, local levels working with professional fire organizations to deliver the five E’s of fire prevention; Engineering, Education, Enforcement, Economic Incentive, and Emergency Response.[6]

Fire investigation is a key element in the fire prevention system that generates data to enable the reduction of fire losses.  The investigation of fires provides a means to identify the cause of the fire in order to develop a knowledge base that could enable the elimination of that cause and thus further reduce the losses from fires.  Data such as the room of fire origin, the first item ignited and ultimately the cause of the fire is information that is critical to understanding and reducing the number of fires.  For example, the identification of products that are involved in ignitions due to a design flaw, in many cases result in the U.S. Consumer Product Safety Commission or a manufacturer issuing a product recall.  
In some cases, the fire investigation determines that the fire was intentional.  An intentional fire as defined in the National Fire Incident Reporting System (NFIRS)  are those fires that are deliberately set and include fires that result from deliberate misuse of a heat source, fires of an incendiary nature (arson), as well as controlled burn fires, such as crop clearing, that required fire service intervention.[7]  An incendiary fire, as defined by the National Fire Protection Association (NFPA) Guide for Fire \& Explosion Investigations, commonly referred to as NFPA 921, is “a fire that is deliberately set with the intent to cause a fire to occur in an area where the fire should not be.”[8]

According to the NFPA, there were approximately 282,600 intentionally set fires each year during the period of 2007 through 2011 in the United States. The average annual loss totals for that period were approximately 420 civilian deaths and 1,360 civilian injuries. More than \$1.3 billion dollars (U.S.) was lost per year due to direct property damage. Fires involving structures account for more than 84~\% of the civilian deaths, injuries and property loss from all intentional fires.  Approximately two-thirds of the structure fires occurred in residential occupancies.  95~\% of the civilian deaths and 86~\% of the civilian injuries also occurred in the intentional fires in residential occupancies. [7]

A subset of the intentional fires are considered arson.  Arson as defined by the U.S. Federal Bureau of Investigation (FBI) Uniform Crime Reporting (UCR) Program defines arson as any willful or malicious burning or attempting to burn, with or without intent to defraud, a dwelling house, public building, motor vehicle or aircraft, personal property of another, etc.[9]

The national rate of arson fire cases cleared by arrest or exceptional means is less than 20~\%.  Of those cases approximately 46~\% of the suspects were prosecuted.[7] Arson is a challenging crime to solve because typically there are no witnesses present at the time of the crime.

Fires are investigated in order to determine the ``origin and cause'' of the fire and in cases of arson there is an effort to determine who was responsible for setting the fire.  This investigation process must follow the scientific method as documented in NFPA 921 [8].  There are numerous textbooks on the subject of conducting a fire investigation [10-15]. The NFPA guide and the texts provide information to a fire investigator collecting data from the scene, analyzing the data, and developing a hypothesis about the fire.

@@

\bibliography{example-thesis}
\end{document}

>>>>

\bibliographystyle{unsrt}

@@

Uncontrolled fires result in loss of life, property and can create an adverse economic impact on a community.  Findings from an archaeological site in South Africa, published in 2012, indicate that mankind has productively used fire for approximately 1 million years~\cite{Berna:2012}.  This is 300 thousand years earlier than previously thought.  Despite mankind’s extensive history with fire, according to the World Health Organization, an estimated 265,000 people die each year as a result of burn injuries~\cite{WHO:2014}.  Based on an estimated global population of 7.3 billion people, the death rate due to burns from all sources would be 3.6 per 100,000 populations.  The World Fire Statistics Center collects data from European, North American and a few Asia-Oceania countries.  Within this group of countries the fire deaths per 100,000 population range from 0.02 (Singapore) to 2.03 (Finland)~\cite{Climate:2014}. There are a significant number of factors that determine the variability of fire death rates between nations, socio-economic conditions being a leading factor~\cite{WHO:2014}.

In the United States, the loss of life due to unwanted fire has been decreasing steadily since the 1970’s.  This is neglecting the increased loss of life as a result of the terrorist attack on September 11, 2001.  The estimated number of fire deaths in the U.S. in 1975 was 8,100 or 3.75 deaths per 100,000 population~\cite{America_Burning_Revisited}.  In 2013, there were 3,240 civilian fire fatalities or 1.02 deaths per 100,000 population~\cite{Karter:2014}.  During this same period, the number of fires in the United States has decreased from 2.5 million fires to 1.24 million fires.  There are many reasons that led to this decrease in loss of life due to fire in the United States.  An example would be research and data that enabled the development of product safety standards, improved building codes, improved fire safety systems, and fire prevention programs to name just a few.  None of this happened without the constant effort of agencies at the Federal, state, and, local levels working with professional fire organizations to deliver the five E’s of fire prevention; Engineering, Education, Enforcement, Economic Incentive, and Emergency Response~\cite{FIVE}.

Fire investigation is a key element in the fire prevention system that generates data to enable the reduction of fire losses.  The investigation of fires provides a means to identify the cause of the fire in order to develop a knowledge base that could enable the elimination of that cause and thus further reduce the losses from fires.  Data such as the room of fire origin, the first item ignited and ultimately the cause of the fire is information that is critical to understanding and reducing the number of fires.  For example, the identification of products that are involved in ignitions due to a design flaw, in many cases result in the U.S. Consumer Product Safety Commission or a manufacturer issuing a product recall.
In some cases, the fire investigation determines that the fire was intentional.  An intentional fire as defined in the National Fire Incident Reporting System (NFIRS)  are those fires that are deliberately set and include fires that result from deliberate misuse of a heat source, fires of an incendiary nature (arson), as well as controlled burn fires, such as crop clearing, that required fire service intervention~\cite{Campbell:2014}.  An incendiary fire, as defined by the National Fire Protection Association (NFPA) Guide for Fire \& Explosion Investigations, commonly referred to as NFPA~921, is ``a fire that is deliberately set with the intent to cause a fire to occur in an area where the fire should not be.''~\cite{NFPA_921}.

According to the NFPA, there were approximately 282,600 intentionally set fires each year during the period of 2007 through 2011 in the United States. The average annual loss totals for that period were approximately 420 civilian deaths and 1,360 civilian injuries. More than \$1.3 billion dollars (U.S.) was lost per year due to direct property damage. Fires involving structures account for more than 84~\% of the civilian deaths, injuries and property loss from all intentional fires.  Approximately two-thirds of the structure fires occurred in residential occupancies.  95~\% of the civilian deaths and 86~\% of the civilian injuries also occurred in the intentional fires in residential occupancies~\cite{Campbell:2014}.

A subset of the intentional fires are considered arson.  Arson as defined by the U.S. Federal Bureau of Investigation (FBI) Uniform Crime Reporting (UCR) Program defines arson as any willful or malicious burning or attempting to burn, with or without intent to defraud, a dwelling house, public building, motor vehicle or aircraft, personal property of another, etc.~\cite{Crime}.

The national rate of arson fire cases cleared by arrest or exceptional means is less than 20~\%.  Of those cases approximately 46~\% of the suspects were prosecuted~\cite{Campbell:2014}. Arson is a challenging crime to solve because typically there are no witnesses present at the time of the crime.

Fires are investigated in order to determine the ``origin and cause'' of the fire and in cases of arson there is an effort to determine who was responsible for setting the fire.  This investigation process must follow the scientific method as documented in NFPA 921~\cite{NFPA_921}.  There are numerous textbooks on the subject of conducting a fire investigation~\cite{Almirall:2004,Fire_Investigation,DeHaan:2012,Icove:2013,Lentinin:2006,Noon:1995}. The NFPA guide and the texts provide information to a fire investigator collecting data from the scene, analyzing the data, and developing a hypothesis about the fire.

@@

\bibliography{../References/Madrzykowski_Thesis}
\end{document}
